Analyzing Performance Bottleneck in Graph Neural Network Training: An Experimental View

### 2020-07-23 Manuscript/sec1-introduction.tex
update by www.grammarly.com

In recent years, the graph neural network (GNN) is a hot research topic in the field of artificial intelligence.
Many GNNs \cite{kipf2017_gcn, defferrad2016_chebnet, li2018_agcn,li2015_ggnn, hamilton2017_graphsage, huang2018_gat, zhang2018_gaan} are proposed.
GNNs can learn the representation of vertices/edges in a graph from its topology and original feature vectors in an \emph{end-to-end} manner.
The strong expressive power makes GNNs achieve impressive accuracy in node classification, link prediction and graph classification \cite{zhou2018_gnn_review, zhang2018_gnn_survey, comprehensive-survey-wu-2020}.

To train GNNs easily, a series of graph neural network learning libraries/systems \cite{PyG, DGL, ma2019_neugraph, zhu2019_aligraph, PGL} are proposed.
PyG \cite{PyG}, NeuGraph \cite{ma2019_neugraph}, PGL \cite{PGL} and DGL \cite{DGL} build upon the existing deep learning frameworks (PyG on PyTorch, NeuGraph on Tensorflow, PGL on PaddlePaddle, DGL on multiple backends).
They provide users with a high-level programming model (message-passing for PyG/PGL/DGL and SAGA-NN for NeuGraph) to describe the structure of a GNN.
They take advantage of the common tools provided by the frameworks like the automatic differentiation to simplify the development.
The GNN learning libraries/systems utilize specially optimized CUDA kernels (like kernel fusion \cite{DGL} \cite{ma2019_neugraph}) and other implementation techniques (like 2D graph partitioning \cite{ma2019_neugraph}) to improve the speed of GNN training.

@@

\begin{itemize}
    \item The training time and the memory usage of a GNN layer is mainly affected by the dimensions of the input/output hidden feature vectors.
          The training time and memory usage increase \emph{linearly} with hyper-parameters.
    \item The edge-related calculation is the performance bottleneck for most GNNs.
          The majority of the training time is spent on it.
          For GNNs with high edge calculation complexity, optimizations should focus on the implementation of the message function that is conducted on every edge during the edge calculation.
          For GNNs with low edge calculation complexity, optimizations should focus on the collect step and the aggregation step of the edge calculation.
          Overlapping the data accessing with the computation is a potential optimizing direction.
    \item The high memory usage caused by the intermediate results of the edge calculation is the main factor that prevents us training big graphs on GPU.
    \item The sampling techniques can significantly reduce the training time and memory usage, but the implementation is still inefficient. The time spent on the sampling may exceed the time spent on the training under big batch sizes. How to make full use of the computing power of GPUs with the sampling techniques is another question to solve.
\end{itemize}

>>>>

In recent years, the graph neural network (GNN) is a hot research topic in the field of artificial intelligence.
Many GNNs \cite{kipf2017_gcn, defferrad2016_chebnet, li2018_agcn,li2015_ggnn, hamilton2017_graphsage, huang2018_gat, zhang2018_gaan} are proposed.
GNN can learn the representation of vertices/edges in a graph from its topology and original feature vectors in an \emph{end-to-end} manner.
The strong expressive power makes GNN achieve impressive accuracy in node classification, link prediction, and graph classification \cite{zhou2018_gnn_review, zhang2018_gnn_survey, comprehensive-survey-wu-2020}.

To train GNN easily, a series of graph neural network learning libraries/systems \cite{PyG, DGL, ma2019_neugraph, zhu2019_aligraph, PGL} are proposed.
PyG \cite{PyG}, NeuGraph \cite{ma2019_neugraph}, PGL \cite{PGL} and DGL \cite{DGL} build upon the existing deep learning frameworks (PyG on PyTorch, NeuGraph on Tensorflow, PGL on PaddlePaddle, DGL on multiple backends).
They provide users with a high-level programming model (message-passing for PyG/PGL/DGL and SAGA-NN for NeuGraph) to describe the structure of a GNN.
They take advantage of the common tools provided by the frameworks like the automatic differentiation to simplify the development.
The GNN learning libraries/systems utilize specially optimized CUDA kernels (like kernel fusion \cite{DGL} \cite{ma2019_neugraph}) and other implementation techniques (like 2D graph partitioning \cite{ma2019_neugraph}) to improve the speed of GNN training.

@@

\begin{itemize}
    \item The training time and the memory usage of a GNN layer is mainly affected by the dimensions of the input/output hidden feature vectors.
          The training time and memory usage increase \emph{linearly} with hyper-parameters.
    \item The edge-related calculation is the performance bottleneck for most GNNs.
          The majority of the training time is spent on it.
          For GNNs with high edge calculation complexity, optimizations should focus on the implementation of the message function that is conducted on every edge during the edge calculation.
          For GNNs with low edge calculation complexity, optimizations should focus on the collect step and the aggregation step of the edge calculation.
          Overlapping the data accessing with the computation is a potential optimizing direction.
    \item The high memory usage caused by the intermediate results of the edge calculation is the main factor that prevents us from training big graphs on GPU.
    \item The sampling techniques can significantly reduce the training time and memory usage, but the implementation is still inefficient. The time spent on the sampling may exceed the time spent on the training under big batch sizes. How to make full use of the computing power of GPUs with the sampling techniques is another question to solve.
\end{itemize}

### 2020-07-23 Manuscript/sec2-review-of-graph-neural-network.tex
update by www.grammarly.com

In this section, we formally define the graph neural network (GNN, for short) and breifly survey typical graph neural networks.
We denote a simple graph $\mathcal{G}$ as $\mathcal{G}=(\mathcal{V}, \mathcal{E})$, where $\mathcal{V}$ and $\mathcal{E}$ are the vertex set and the edge set of $\mathcal{G}$, respectively.
Let $n=|\mathcal{V}|$ and $m=|\mathcal{E}|$ as the number of vertices/edges.
We use $v_i$ $(0 \leq i < n)$ to denote a vertex and $e_{i,j}=(v_i, v_j)$ to denote the edge pointing from $v_i$ to $v_j$.
The adjacency set of $v_i$ is $\mathcal{N}(v_i)=\{v|(v_i, v) \in \mathcal{E}\}$.
We denote a \emph{vector} with a bold lower case letter like $\boldsymbol{x}$ and a \emph{matri}x with a bold upper case letter like $\boldsymbol{X}$.

@@

A GNN layer consists of $n$ graph neurons, where $n$ is the number of vertices in $\mathcal{G}$.
Each graph neuron corresponds to a vertex in $\mathcal{G}$.
In the first GNN layer (Layer 0), the graph neuron of the vertex $v_i$ collects input feature vectors of itself and the vertices $\boldsymbol{x}_j$ that are adjacent to $v_i$ in $\mathcal{G}$ (i.e., $v_j \in \mathcal{N}(v_i)$) from the input layer.
After aggregating input feature vectors and applying non-linear transformation, the graph neuron outputs a hidden feature vector $\boldsymbol{h}^1_i$ for $v_i$.
Take the demo DNN in \figurename~\ref{fig:general_structure_of_gnn}(a) as the example.
Since $\mathcal{N}(v_3) = \{v_1, v_2, v_4, v_5, v_6\}$, the graph neuron of $v_1$ at layer 0 collects the feature vectors \{$\boldsymbol{x}_1$, $\boldsymbol{x}_2$, $\boldsymbol{x}_3$, $\boldsymbol{x}_4$, $\boldsymbol{x}_5$, $\boldsymbol{x}_6$\} from the input layer and outputs $\boldsymbol{h}^1_1$.
Different GNNs mainly differ in the graph neurons that they use.
We elaborate on their details later.

@@

In the next GNN layer (Layer 1), the graph neuron of $v_i$ collects the hidden feature vectors of itself $\boldsymbol{h}^1_i$ and its neighbors ($\boldsymbol{h}^1_j$ with $v_j \in \mathcal{N}(v_i)$) from the \emph{previous} GNN layer.
Based on the collected hidden vectors, the graph neuron in Layer 1 outputs a new hidden feature vector $\boldsymbol{h}^2_i$ for $v_i$.
Though there are only two GNN layers in \figurename~\ref{fig:general_structure_of_gnn}, a GNN allows to stack more GNN layers to support deeper graph analysis.
%In \figurename~\ref{fig:general_structure_of_gnn}, we shows a GNN with two layers.

Assume there are $L$ GNN layers.
The last GNN layer (Layer $L-1$) outputs a hidden feature vector $\boldsymbol{h}^{L}_i$ for every vertex $v_i$.
As an embedding vector, $\boldsymbol{h}^L_i$ encodoes the knowledge learned from the input layer and all the previous GNN layers.
Since $\boldsymbol{h}^L_i$ is affected by $v_i$ and the vertices in the $L$-hop neighborhood of $v_i$, analyzing a graph with a \emph{deeper} GNN means analyzing each vertex with a \emph{wider} scope.

The hidden feature vectors $\boldsymbol{h}^L_i$ of the last GNN layer are fed to the prediction layer to generate the output of the whole GNN.
The prediction layer is a standard nerual network.
The structure of the prediction layer depends on the prediction task of the GNN.
Take the node classification task as the example, as shown in \figurename~\ref{fig:general_structure_of_gnn}.
The node classification predicts a label for every vertex in $\mathcal{G}$.
In this case, the prediction layer can be a simple softmax layer with $\boldsymbol{h}^L_i$ as the input and a vector of probabilities as the output.
If the prediction task is edge prediction, the hidden feature vectors of two vertices are concatenated and fed into a softmax layer.
If we need to predict a label for the whole graph, a pooling (max/mean/...) layer is added to generate an embedding vector for the whole graph and the embedding vector is used to produce the final prediction.

Supporting end-to-end training is a prominent advantage of GNN, compared with other graph-based machine learning methods.
We can calculate the gradients of the loss function on the model parameters from the prediction layer directly.
With the help of the back proporgation technique, the gradient is propogated from the prediction layer back to the previous GNN layers layer by layer.
The model parameters are updated with a gradient descent optimizer like Adam.
Except for the input feature vector, there is no need to conduct handworked feature extraction.
In a fully parameterized way, the GNN automatically extracts an embedding vector for each vertex from its $L$-hop neighborhood.
The parameters are tuned according to the specific prediction task, leading to high prediction accuracy.

@@

In \tablename~\ref{tab:gnn_overview_edge} and \tablename~\ref{tab:gnn_overview_vertex}, we list the edge and vertex calculation complexity, respectively.
The time complexity of a graph neuron are affected by the dimensions of the input/output hidden vectors $d_{in}$ and $d_{out}$ and the dimensions of the model parameters (like the number of heads $K$ in GAT and the dimensions of the view vectors $d_a$/$d_v$ in GaAN).

We classify the typical GNNs into four quadrants based on their edge/vertex complexity as shown in \figurename~\ref{fig:gnn_complexity_quadrant}. We pick GCN, GGNN, GAT and GaAN as the \emph{representative GNNs} of the four quadrants.

@@

\textbf{GGNN} (high vertex \& low edge calculation complexity):
GGNN introduces the gated recurrent unit (GRU) into the graph neural networks.
The vertex update function $\phi$ of GGNN is a modified GRU unit that has 12 model parameters to learn, having high computational complexity.
To lower the training cost, all GNN layers share the same group of parameters in GGNN.
GGNN further requires the dimension of $\boldsymbol{h}^{l+1}$ is equal to the dimension of $\boldsymbol{h}^l$.
Since the message function $\phi$ only uses the hidden feature vector $\boldsymbol{h}^l_j$ of the source vertex $v_j$ of an edge $(v_j, v_i)$, in the implementation, GGNN conducts the pre-processing vertex calculation $\hat{\boldsymbol{h}}^l_i=\boldsymbol{W}\boldsymbol{h}^l_i$ for every vertex $v_i$ before the message-passing.
The message function $\phi$ directly uses $\hat{\boldsymbol{h}}^l_j$ as the message vector for the edge $(v_j, v_i)$.
By this way, GGNN further reduces the time complexity of the edge calculation to $O(1)$ without increasing the time complexity of the vertex calculation.

\textbf{GAT} (low vertex \& high edge calculation complexity):
GAT introduces the attention and multi-head mechanism into the graph neural networks.
The $K$ heads generate $K$ independent views for an edge, where $K$ is a hyper-parameter.
The views of $K$ heads can be merged by concatenating or by averaging.
For concatenating, the dimension of the hidden feature vector of each head $d_{head}$ is $d_{out}/K$.
For averaging, $d_{head}$ is $d_{out}$, multiplying the complexity by $K$.
Each GAT layer consists of a preprocessing step and two sub-layers.
In the preprocessing step, GAT calculates the attention vectors $\hat{\boldsymbol{h}}^{l}_{i,(k)}$ for every vertex $v_i$ in every head $k$.
The first sub-layer uses the attention vectors to calculate the attention weights of every edge in every head $\boldsymbol{m}^{l,0}_{j,i,(k)}$.
After the message-passing, the first-layer gets the summation of attention weights of every vertex in every head $\hat{\boldsymbol{a}}^{l,0}_{i,(k)}$.
The second sub-layer gets the normalized attention weights $\alpha_{j, i, (k)}$ for every edge in every head and aggregates the hidden feature vectors with the normalized weights in the message-passing.
GAT uses the aggreated vector $\boldsymbol{s}^{l,1}_i$ of the second sub-layer as the output hidden vector directly.

\textbf{GaAN} (high vertex \& edge calculation complexity):
Based on the multi-head machanism, GaAN introduces a convolutional subnetwork to control the weight of each head. Different to multi-head machanism in GAT, GaAN project the center vertex feature to the query vector (dimension is $d_v$) and
the neighboring vertex features are projected to get key and value vectors. The dimension of key and values vectors are the same, we set to $d_a$.
$K$ is the number of attention heads. The weight of each head is calculated by a convolutional network, which takes the feature of each vertex and its neighbors as input to get the gated values with number of $K$.
The convolutional network is built by combining average pooling and max pooling. Moreover, the max pooling needs to project the vertex feature to a transformed vector with dimension of $d_m$, costing an additional computational overhead.
Each GaAN layer consists of four sub-layers. The first sub-layer aims to get the summation of attention weights of every vertex in each head.
The second sub-layer uses to aggreate neighboring vertices's query vector with the normalized attention weights $\alpha_{j, i, (k)}$ for every head.
And the third sub-layer is used in max-pooling and it gets the element-wise max vector of neighboring vertices' transformed vectors by message-passing.
The last sub-layer gets element-wise mean vector of neighboring vertices' vectors by message-passing. The gate values $\boldsymbol{g}^l_i$ is obtained by using a fully-connected layer with vector concatenating $\boldsymbol{h}^{l}_i$, $\boldsymbol{s}^{l, 2}_i$ and $\boldsymbol{s}^{l, 3}_i$.
At last, GaAN uses a fully-connected layer with concatenating $\boldsymbol{h}^{l}_i$ and the result of element-wise multiplying $\boldsymbol{g}^{l}_i$ and $\boldsymbol{s}^{l, 0}_i$ to get the output hidden vector.

@@

To train GNNs in a mini-batch way, several sampling techniques are proposed.
In each mini-batch, they sample a small subgraph from the whole graph $\mathcal{G}$ and uses the subgraph to update the model parameters.
In other words, the sampling techniques only active the graph neurons and the connections that appear in the sampled subgraph between GNN layers, as shown in \figurename~\ref{fig:gnn_sampling}.
Inactive graph neurons and connetions do not participate in the training of this mini-batch, saving lots of computation and storage costs.
Moreover, it may reduce the risk of overfitting the training graph.
Based on whether different GNN layers sample different subgraphs, the existing sampling techniques can be devided into two groups \cite{zeng2020_graphsaint}: the neighbor sampling and the graph sampling.

The neighbor sampling techniques \cite{hamilton2017_graphsage, ying2018_pinsage, chen2018_fastgcn, chen2018_sgcn, huang2018_adap} sample the subgraphs layer by layer.
They first sample several vertices from $\mathcal{V}$ in the last GNN layer.
Then they repeatedly sample the neighbors of those vertices in the previous layer until the input layer.
The sampled subgraphs of different GNN layers may be \emph{different}, as shown in \figurename~\ref{fig:gnn_sampling_neighbor_sampling}.
GraphSAGE \cite{hamilton2017_graphsage} is the representative technique.
For every sampled vertex $v_i$ in the GNN layer $l$, GraphSAGE samples at most $S^l$ neighbors of $v_i$ in the previous GNN layer.
$S^l$ is the hyperparameter that is usually much smaller than $|\mathcal{V}|$.
By this way, GraphSAGE limits the neighborhood sizes of the vertices in the sampled subgraph, especially high-degree vertices.

>>>>

In this section, we formally define the graph neural network (GNN, for short) and briefly survey typical graph neural networks.
We denote a simple graph $\mathcal{G}$ as $\mathcal{G}=(\mathcal{V}, \mathcal{E})$, where $\mathcal{V}$ and $\mathcal{E}$ are the vertex set and the edge set of $\mathcal{G}$, respectively.
Let $n=|\mathcal{V}|$ and $m=|\mathcal{E}|$ as the number of vertices/edges.
We use $v_i$ $(0 \leq i < n)$ to denote a vertex and $e_{i,j}=(v_i, v_j)$ to denote the edge pointing from $v_i$ to $v_j$.
The adjacency set of $v_i$ is $\mathcal{N}(v_i)=\{v|(v_i, v) \in \mathcal{E}\}$.
We denote a \emph{vector} with a bold lower case letter like $\boldsymbol{x}$ and a \emph{matri}x with a bold upper case letter like $\boldsymbol{X}$.

@@

A GNN layer consists of $n$ graph neurons, where $n$ is the number of vertices in $\mathcal{G}$.
Each graph neuron corresponds to a vertex in $\mathcal{G}$.
In the first GNN layer (Layer 0), the graph neuron of the vertex $v_i$ collects input feature vectors of itself and the vertices $\boldsymbol{x}_j$ that are adjacent to $v_i$ in $\mathcal{G}$ (i.e., $v_j \in \mathcal{N}(v_i)$) from the input layer.
After aggregating input feature vectors and applying the non-linear transformation, the graph neuron outputs a hidden feature vector $\boldsymbol{h}^1_i$ for $v_i$.
Take the demo DNN in \figurename~\ref{fig:general_structure_of_gnn}(a) as the example.
Since $\mathcal{N}(v_3) = \{v_1, v_2, v_4, v_5, v_6\}$, the graph neuron of $v_1$ at layer 0 collects the feature vectors \{$\boldsymbol{x}_1$, $\boldsymbol{x}_2$, $\boldsymbol{x}_3$, $\boldsymbol{x}_4$, $\boldsymbol{x}_5$, $\boldsymbol{x}_6$\} from the input layer and outputs $\boldsymbol{h}^1_1$.
Different GNNs mainly differ in the graph neurons that they use.
We elaborate on their details later.

@@

In the next GNN layer (Layer 1), the graph neuron of $v_i$ collects the hidden feature vectors of itself $\boldsymbol{h}^1_i$ and its neighbors ($\boldsymbol{h}^1_j$ with $v_j \in \mathcal{N}(v_i)$) from the \emph{previous} GNN layer.
Based on the collected hidden vectors, the graph neuron in Layer 1 outputs a new hidden feature vector $\boldsymbol{h}^2_i$ for $v_i$.
Though there are only two GNN layers in \figurename~\ref{fig:general_structure_of_gnn}, a GNN allows us to stack more GNN layers to support deeper graph analysis.
%In \figurename~\ref{fig:general_structure_of_gnn}, we shows a GNN with two layers.

Assume there are $L$ GNN layers.
The last GNN layer (Layer $L-1$) outputs a hidden feature vector $\boldsymbol{h}^{L}_i$ for every vertex $v_i$.
As an embedding vector, $\boldsymbol{h}^L_i$ encodes the knowledge learned from the input layer and all the previous GNN layers.
Since $\boldsymbol{h}^L_i$ is affected by $v_i$ and the vertices in the $L$-hop neighborhood of $v_i$, analyzing a graph with a \emph{deeper} GNN means analyzing each vertex with a \emph{wider} scope.

The hidden feature vectors $\boldsymbol{h}^L_i$ of the last GNN layer are fed to the prediction layer to generate the output of the whole GNN.
The prediction layer is a standard neural network.
The structure of the prediction layer depends on the prediction task of the GNN.
Take the node classification task as the example, as shown in \figurename~\ref{fig:general_structure_of_gnn}.
The node classification predicts a label for every vertex in $\mathcal{G}$.
In this case, the prediction layer can be a simple softmax layer with $\boldsymbol{h}^L_i$ as the input and a vector of probabilities as the output.
If the prediction task is edge prediction, the hidden feature vectors of two vertices are concatenated and fed into a softmax layer.
If we need to predict a label for the whole graph, a pooling (max/mean/...) layer is added to generate an embedding vector for the whole graph and the embedding vector is used to produce the final prediction.

Supporting end-to-end training is a prominent advantage of GNN, compared with other graph-based machine learning methods.
We can calculate the gradients of the loss function on the model parameters from the prediction layer directly.
With the help of the backpropagation technique, the gradient is propagated from the prediction layer back to the previous GNN layers layer by layer.
The model parameters are updated with a gradient descent optimizer like Adam.
Except for the input feature vector, there is no need to conduct handworked feature extraction.
In a fully parameterized way, the GNN automatically extracts an embedding vector for each vertex from its $L$-hop neighborhood.
The parameters are tuned according to the specific prediction task, leading to high prediction accuracy.

@@

In \tablename~\ref{tab:gnn_overview_edge} and \tablename~\ref{tab:gnn_overview_vertex}, we list the edge and vertex calculation complexity, respectively.
The time complexity of a graph neuron is affected by the dimensions of the input/output hidden vectors $d_{in}$ and $d_{out}$ and the dimensions of the model parameters (like the number of heads $K$ in GAT and the dimensions of the view vectors $d_a$/$d_v$ in GaAN).

We classify the typical GNNs into four quadrants based on their edge/vertex complexity as shown in \figurename~\ref{fig:gnn_complexity_quadrant}. We pick GCN, GGNN, GAT, and GaAN as the \emph{representative GNNs} of the four quadrants.

@@

\textbf{GGNN} (high vertex \& low edge calculation complexity):
GGNN introduces the gated recurrent unit (GRU) into the graph neural networks.
The vertex update function $\phi$ of GGNN is a modified GRU unit that has 12 model parameters to learn, having high computational complexity.
To lower the training cost, all GNN layers share the same group of parameters in GGNN.
GGNN further requires the dimension of $\boldsymbol{h}^{l+1}$ is equal to the dimension of $\boldsymbol{h}^l$.
Since the message function $\phi$ only uses the hidden feature vector $\boldsymbol{h}^l_j$ of the source vertex $v_j$ of an edge $(v_j, v_i)$, in the implementation, GGNN conducts the pre-processing vertex calculation $\hat{\boldsymbol{h}}^l_i=\boldsymbol{W}\boldsymbol{h}^l_i$ for every vertex $v_i$ before the message-passing.
The message function $\phi$ directly uses $\hat{\boldsymbol{h}}^l_j$ as the message vector for the edge $(v_j, v_i)$.
In this way, GGNN further reduces the time complexity of the edge calculation to $O(1)$ without increasing the time complexity of the vertex calculation.

\textbf{GAT} (low vertex \& high edge calculation complexity):
GAT introduces the attention and multi-head mechanism into the graph neural networks.
The $K$ heads generate $K$ independent views for an edge, where $K$ is a hyper-parameter.
The views of $K$ heads can be merged by concatenating or by averaging.
For concatenating, the dimension of the hidden feature vector of each head $d_{head}$ is $d_{out}/K$.
For averaging, $d_{head}$ is $d_{out}$, multiplying the complexity by $K$.
Each GAT layer consists of a preprocessing step and two sub-layers.
In the preprocessing step, GAT calculates the attention vectors $\hat{\boldsymbol{h}}^{l}_{i,(k)}$ for every vertex $v_i$ in every head $k$.
The first sub-layer uses the attention vectors to calculate the attention weights of every edge in every head $\boldsymbol{m}^{l,0}_{j,i,(k)}$.
After the message-passing, the first-layer gets the summation of attention weights of every vertex in every head $\hat{\boldsymbol{a}}^{l,0}_{i,(k)}$.
The second sub-layer gets the normalized attention weights $\alpha_{j, i, (k)}$ for every edge in every head and aggregates the hidden feature vectors with the normalized weights in the message-passing.
GAT uses the aggregated vector $\boldsymbol{s}^{l,1}_i$ of the second sub-layer as the output hidden vector directly.

\textbf{GaAN} (high vertex \& edge calculation complexity):
Based on the multi-head mechanism, GaAN introduces a convolutional subnetwork to control the weight of each head. Different from to multi-head mechanism in GAT, GaAN project the center vertex feature to the query vector (dimension is $d_v$), and
the neighboring vertex features are projected to get key and value vectors. The dimension of key and values vectors are the same, we set to $d_a$.
$K$ is the number of attention heads. The weight of each head is calculated by a convolutional network, which takes the feature of each vertex and its neighbors as input to get the gated values with a dimension of $K$.
The convolutional network is built by combining average pooling and max pooling. Moreover, the max-pooling needs to project the vertex feature to a transformed vector with a dimension of $d_m$, costing an additional computational overhead.
Each GaAN layer consists of four sub-layers. The first sub-layer aims to get the summation of attention weights of every vertex in each head.
The second sub-layer uses to aggregate neighboring vertices' query vector with the normalized attention weights $\alpha_{j, i, (k)}$ for every head.
And the third sub-layer is used in max-pooling and it gets the element-wise max vector of neighboring vertices' transformed vectors by message-passing.
The last sub-layer gets element-wise mean vector of neighboring vertices' vectors by message-passing. The gate values $\boldsymbol{g}^l_i$ is obtained by using a fully-connected layer with vector concatenating $\boldsymbol{h}^{l}_i$, $\boldsymbol{s}^{l, 2}_i$ and $\boldsymbol{s}^{l, 3}_i$.
At last, GaAN uses a fully-connected layer with concatenating $\boldsymbol{h}^{l}_i$ and the result of element-wise multiplying $\boldsymbol{g}^{l}_i$ and $\boldsymbol{s}^{l, 0}_i$ to get the output hidden vector.

@@

To train GNNs in a mini-batch way, several sampling techniques are proposed.
In each mini-batch, they sample a small subgraph from the whole graph $\mathcal{G}$ and uses the subgraph to update the model parameters.
In other words, the sampling techniques only active the graph neurons and the connections that appear in the sampled subgraph between GNN layers, as shown in \figurename~\ref{fig:gnn_sampling}.
Inactive graph neurons and connections do not participate in the training of this mini-batch, saving lots of computation and storage costs.
Moreover, it may reduce the risk of overfitting the training graph.
Based on whether different GNN layers sample different subgraphs, the existing sampling techniques can be divided into two groups \cite{zeng2020_graphsaint}: the neighbor sampling and the graph sampling.

The neighbor sampling techniques \cite{hamilton2017_graphsage, ying2018_pinsage, chen2018_fastgcn, chen2018_sgcn, huang2018_adap} sample the subgraphs layer by layer.
The first sample several vertices from $\mathcal{V}$ in the last GNN layer.
Then they repeatedly sample the neighbors of those vertices in the previous layer until the input layer.
The sampled subgraphs of different GNN layers may be \emph{different}, as shown in \figurename~\ref{fig:gnn_sampling_neighbor_sampling}.
GraphSAGE \cite{hamilton2017_graphsage} is the representative technique.
For every sampled vertex $v_i$ in the GNN layer $l$, GraphSAGE samples at most $S^l$ neighbors of $v_i$ in the previous GNN layer.
$S^l$ is the hyperparameter that is usually much smaller than $|\mathcal{V}|$.
In this way, GraphSAGE limits the neighborhood sizes of the vertices in the sampled subgraph, especially high-degree vertices.
